Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-15 qpip0.tex
Write proof first draft for Pi_Int binding

	Protocol:
	\begin{enumerate}
		\item \label{step:naive1} The verifier chooses a $XZ$-measurement $h$ from the distribution specified in \stepref{qpip1-verify} of $\PiSamp$.
		\item \label{step:naive2} The prover prepares $\rho$ by running \stepref{qpip1-state-gen} of $\PiSamp$.
		\item \label{step:urmila-in-naive}
		    The verifier and prover run $(\PMeasure(\rho), \VMeasure(h))(1^\lambda)$. On $\PiMeasure$'s second message, we allow the prover to abort by sending $a=\bot$, which does not benefit cheating provers but is useful for analysis.

			If it's a testing round, the verifier accepts or rejects based on the outcome of $\PiMeasure$.
			If it's a Hadamard round, the verifier obtains $v$.
		\item \label{step:naive-output} If it's a Hadamard round, the verifier finishes the verification step of Protocol~\ref{ProtoQPIP1} by generating and outputting $(d, z)$

	\end{enumerate}
\end{protocol}

@@

\Ethan{TODO move the following}

\begin{lem}[another triangle inequality for computational distance]\label{lem:computational-triangle}
Let $D_1$, $D_2$ be two distributions such that $D_1=D_3+D_4$ and $D_2=D_5+D_6$, where $D_3$, $D_4$, $D_5$, and $D_6$ are subnormalized distributions.

If
\begin{align}
	D_1\approx_{c,\eps} D_2 \\
	D_4\approx_{c,\eps'} D_6,
\end{align}
then
\begin{align}
	D_3\approx_{c,\eps+\eps'} D_5
\end{align}
\end{lem}
% \begin{align}
%     D_1 &\approx_{q,\eps} D_2 \\
%     D_4 &\approx_{q,\eps'} D_6,
% \end{align}
% then
% \begin{align}
%     D_3 \approx_{q,\eps+\eps'} D_5
% \end{align}
\begin{proof}
% Since
% $$\norm{D_4 - D_6}_q \leq \eps',$$
% there exist

We prove by contradiction. Suppose that
$$D_3\not\approx_{c, \eps+\eps'} D_5,$$
then there exist a quantum circuit $A$ of size $poly(n)$ such that
$$\L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1]  \R\| > \eps+\eps'.$$
Since $$D_4\approx_{c, \eps'}D_6,$$ we know that
$$\L\|\Pr[A({D_4})=1]-\Pr[A({D_6})=1]  \R\| \leq \eps'.$$
Since $D_1=D_3+D_4$ and $D_2=D_5+D_6$, we have that
$$ \Pr[A({D_1})=1]= \Pr[A({D_3})=1]+ \Pr[A({D_4})=1]$$
$$ \Pr[A({D_2})=1]= \Pr[A({D_5})=1]+ \Pr[A({D_6})=1].$$
So
\begin{align}
    &\L\|\Pr[A({D_1})=1]-\Pr[A({D_2})=1]  \R\| \nn \\
    =&   \L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1] +\Pr[A({D_4})=1]-\Pr[A({D_6})=1]  \R\| \nn \\
    \geq  &\L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1]  \R\| - \L\|\Pr[A({D_6})=1]-\Pr[A({D_4})=1]  \R\|  \nn \\
     > &\eps
\end{align}
	This contradicts with the assumption that $D_1\approx_{c,\eps}D_2$.
\end{proof}

\hannote{comp. indistinguish for subnormalized state?}
\hannote{triangle inequality}

\begin{lemma}[binding property of $\PiNaive$]
\label{lem:naive-qpip0-binding}
	Let $\PNaiveStar$ be a cheating $\BQP$ prover for $\PiNaive$ and $\lambda$ be the security parameter.
	Under the QLWE assumption, if conditioned on $\PNaiveStar$ not aborting, $\PNaiveStar$ passes the testing round with probability $1-\negl(\lambda)$,
	\Ethan{Wrong assumption. The ``perfect prover" notion here is generalized and allows giving up.}
	then the verifier's output in the Hadamard round is $(\eps+\negl(\lambda))$-computationally indistinguishable from $(d, z_{ideal})$.
\end{lemma}

\begin{proof}
	\Ethan{TODO this is currently wrong and I'm not sure how to fix it.
	The version before this which had subnormalized states seems strange too.
	The correct proof should involve an alternative prover strategy that ``fixes" the abort case
	by re-running with perfect prover strategy,
	in order to connect the dots with \Cref{lem:urmila-binding}}
	By \Cref{lem:urmila-binding},
	there exist some $\rho$ such that
	$v\approx_{c,\negl(\lambda)}M_{XZ}(\rho,h)$.
	Let $\cVVer$ be the verification operation in \stepref{qpip1-verify} of $\PiSamp$.
	Since $\cVVer$ is efficient, we have
	$(d,z)\approx_{c, \negl(\lambda)}\cVVer(M_{XZ}(\rho, h))$.
	\Ethan{There were some shenanigans involving subnormalized states for which I'm not quite sure what's going on.}
	By \Cref{QPIP1thm}, for all $\rho$, 
	$\cVVer(M_{XZ}(\rho, h))\approx_{c, \eps}(d, z_{ideal})$.
	So we conclude that
	$(d, z)\approx_{c, \eps+\negl(\lambda)}(d, z_{ideal})$.

	\Ethan{Original proof below.}

	By Lemma~\ref{lem:urmila-binding}, if the prover passes the testing round with probability $1-\negl(n)$, there exist a $\rho$ such that $\norm{z_{commit}- M_{XZ}(\rho,h)}_q \leq \negl(n)$. Note that the prover can choose to abort, so $z_{commit}$ and $\rho$ are subnormalized. Let $Ver$ be the verification operation run by Protocol~\ref{ProtoQPIP1} on the measurement outcome. \hannote{describe better} Since $Ver$ is an efficient process and $Ver(z_{commmit})+p_{abort}(\Rej,\bot)=(d,z_{naive})$, we have that

\begin{equation} \label{eq:z=verM}
    \norm{(d,z_{naive})-(Ver( M_{XZ}(\rho,h))+p_{abort}(\Rej,\bot))}_q \leq \negl(n).
\end{equation}
\hannote{actually true with subnormalize shit?}
 By Theorem~\ref{QPIP1thm}, for all $\rho$,
 \begin{equation}
     \norm{Ver(M_{XZ}(\rho,h))-(d,z_{ideal})} \leq O(\eps).
 \end{equation}
   Combined with Equation~\ref{eq:z=verM} and Lemma~\ref{lem:computational-triangle0}, and note that $\norm{(\Rej,\bot)-(d,z_{ideal})}=0$, we have that
   \begin{equation}
       \norm{(d,z_{naive})-(d,z_{ideal})}_q\leq O(\eps)
   \end{equation}
\end{proof}

\begin{proof}
For every cheating prover $\PNaiveStar$ which aborts, consider an alternative prover $P_{alt}$ which is $\PNaiveStar$ with the following modification: on the second round of $\PiMeasure$, if $\PNaiveStar$ would abort, $P_{alt}$ switch to run the trivial strategy specified by Fact~\ref{fact:perfectly_pass_test} instead. This is possible because the message send by the verifier on the first round is classical, so the prover can recalculate everything.

By \Cref{lem:trivial-4-round-strategy} and the assumption of the lemma, $p_{alt}$ passes the testing round with $(1-\negl(\lambda))$ probability. Therefore by Lemma~\ref{lem:urmila-binding}, there exist a state $\rho_{alt}$ such that 
\begin{align} \label{eq:z-alt-rho}
    z_{commit,alt} \approx_{c,\negl(\lambda)} M_{XZ}(\rho_{alt},h).
\end{align}

Note that $z_{commit,alt} = z_{commit}+z_{trivial}$, where $z_{commit}$ and $z_{trivial}$ are sub-normalized states correspond to whether $p_{alt}$ follows $\PNaiveStar$ or the trivial strategy.

By Fact~\ref{lem:trivial-4-round-strategy} and Lemma~\ref{lem:urmila-binding}, there exist a sub-normalized state $\rho_{trivial}$ such that

\begin{align} \label{eq:z-triv-rho}
    z_{trivial} \approx_{c,\negl(\lambda)} M_{XZ}(\rho_{trivial},h).
\end{align}

By Equation~\ref{eq:z-alt-rho}, Equation~\ref{eq:z-triv-rho}, and Lemma~\ref{lem:computational-triangle}, we have

\begin{align} \label{eq:z-triv-rho}
    z_{commit} \approx_{c,\negl(\lambda)} M_{XZ}(\rho_{commit},h),
\end{align}

where $\rho_{commit} \defeq \rho_{alt}-\rho_{trivial}$.\hannote{wrong, not a state.} Note that

>>>>

	Protocol:
	\begin{enumerate}
		\item \label{step:naive1} The verifier chooses a $XZ$-measurement $h$ from the distribution specified in \stepref{qpip1-verify} of $\PiSamp$.
		\item \label{step:naive2} The prover prepares $\rho$ by running \stepref{qpip1-state-gen} of $\PiSamp$.
		\item \label{step:urmila-in-naive}
		    The verifier and prover run $(\PMeasure(\rho), \VMeasure(h))(1^\lambda)$.
			\begin{enumerate}
				\item The verifier samples $(pk, sk)\leftarrow\cVNaive{1}(1^\lambda, h)$ and sends $pk$ to the prover.
				\item The prover runs $(y, \sigma)\leftarrow\cPNaive{2}(pk, \rho)$.
					Here we allow the prover to abort by sending $y=\bot$, which does not benefit cheating provers but is useful for analysis.
				\item The verifier samples $c\leftarrow\cVNaive{3}$
				\item The prover replies $a\leftarrow\cPNaive{4}(pk, c, \sigma)$.
				\item 
					If it is a testing round, the verifier accepts or rejects based on the outcome of $\PiMeasure$.
					If it is a Hadamard round, the verifier obtains $v$.
			\end{enumerate}
		\item \label{step:naive-output} If it's a Hadamard round, the verifier finishes the verification step of Protocol~\ref{ProtoQPIP1} by generating and outputting $(d, z)$
	\end{enumerate}
\end{protocol}

@@

\begin{lemma}[binding property of $\PiNaive$]
\label{lem:naive-qpip0-binding}
	Let $\PNaiveStar$ be a cheating $\BQP$ prover for $\PiNaive$ and $\lambda$ be the security parameter.
	Under the QLWE assumption, if conditioned on $\PNaiveStar$ not aborting, $\PNaiveStar$ passes the testing round with probability $1-\negl(\lambda)$,
	then the verifier's output in the Hadamard round is $(\eps+\negl(\lambda))$-computationally indistinguishable from $(d, z_{ideal})$.
\end{lemma}
\begin{proof}
	Consider the following reduction to another cheating $\BQP$ prover for $\PiNaive$ that is perfect and non-aborting.

	We define $\Pstar$ as follows.
	For the second message, run $(y, \sigma)\leftarrow\cPNaiveStar{2}(pk, \rho)$.
	If $y\ne\bot$, then reply $y$.
	Else run the corresponding step of the dummy strategy as in \Cref{lem:trivial-4-round-strategy} and reply with its results.
	For the fourth message, if $y\ne\bot$, run and reply with $a\leftarrow\cPNaiveStar{4}(pk, c, \sigma)$.
	If $y=\bot$ and $c=0$, continue the dummy strategy.

	Observe that this $\Pstar$ indeed passes testing round with overwhelming probability by construction.
	So we can apply \Cref{lem:urmila-binding} to the $\PiMeasure$ call to use its binding property.
	Combining it with $\PiSamp$'s soundness,
	we see that the verifier's outputs corresponding $\Pstar$, $(d', z')$ is computationally indistinguishable to some $(d', z_{ideal}')$

	Conditioned on $\PNaiveStar$ accept, clearly $(d, z)=(d', z')$.
	On $\PNaiveStar$ reject, since dummy strategy will be rejected with overwhelming probability in Hadamard round, \Ethan{kinda obvious but we never explicitly explained why}
	we also have $(d, z)=(\Rej, \bot)\approx_{c, \negl(\lambda)}(d', z')$

	The conclusion follows from a simple hybrid argument.
\end{proof}
